\LARGE \bf Enriching Robot's Actions with Affective Movements

### 2017-03-24 RO-MAN-2017-2/Parts/EmotionalTree.tex
Changes in some of the files. Improvements in explanations that were missing before. Improvement of the EBFN description. Shift of the section description after the introduction of the emotional execution three to improve the readability of the description section. I still need to correct and extend some explanations in the article.

Emotional Execution Tree is a connected acyclic graph $G(V,E)$ with $|V|$ vertexes and $|E|$ edges. The root and non-leaf nodes could be of either \textit{parallel} or \textit{sequential} type. Sequential nodes execute one branch after the other. Once all the branches have been executed, it notifies to its predecessor to continue with the execution of other nodes.  The parallel node offers two type of synchronization. one is at action and the other at emotion. This distinction between action and emotion synchronization enables the possibility to change emotion parameter without affecting the action. done could be one out of four different sub-types: (i) action and emotion synchronous, (ii) action synchronous and emotion asynchronous, (iii) action asynchronous and emotion synchronous, or (iv) action and emotion asynchronous. 
 Action synchronous means that each time that a parallel node receives a ``finish'' notification (i.e., success or failure), it will broadcast the message to all nodes that derived it and to its predecessor, which will stop their execution.

 If a sequence node receives a finish message, it will execute the next branch. When all branches have been executed, it communicates the end of the action. On the other hand, emotion synchronous means that each time that a node (either sequence or parallel) receives an emotion synchronization message, it will propagate the message to all branches to move to the consecutive emotional expression. If a node is principal and it has finished to execute all actions, it will notify its predecessor.

This distinction creates the possibility to synchronize emotional changes without affecting the normal execution of an action and it also enables synchronization among parallel actions. Finally, leaf nodes could only be simple action nodes that have been implemented in the system. Any node can belong to one of two levels: principal or  secondary. If a node is principal, it will notify its predecessor about the messages that it has received, while the secondary node cannot propagate any message to its predecessor. This enables the possibility to have actions that are the ones to command over the others. 
Compound actions are implemented combining all type of nodes.

>>>>

Emotional Execution Tree is a connected acyclic graph $G(V,E)$ with $|V|$ vertexes and $|E|$ edges. The root and non-leaf nodes could be of either \textit{parallel} or \textit{sequential} type, which represent the type of actions that would be executed by the system. Sequential nodes execute one branch after the other. Once all the branches have been executed, it notifies its predecessor to continue with the execution of other branches. The parallel node executes all the branches at the same time. This type of node offers two levels of synchronization: action and emotion. The action level synchronization enables the possibility to stop all the branches once the main branch has completed its execution. This main branch is established in the action message. On the other hand, the emotion level synchronization update to a new emotional parameter of all the branches once the main branch updates to a  new emotional parameter. In both cases, any update or finish message will be ignored if the branch is not principal. These two levels of synchronization creates four sub-types of parallel nodes: (i) action and emotion synchronous, (ii) action synchronous and emotion asynchronous, (iii) action asynchronous and emotion synchronous, or (iv) action and emotion asynchronous.

Finally, leaf nodes are only simple action nodes that have been implemented in the system. Any node can belong to one of two levels: principal or secondary. If a node is principal, it will notify its predecessor about the messages that it has received, while the secondary node cannot propagate any message to its predecessor. This enables the possibility to have actions that have priority over others. To exemplify this, lets continue with the sequence of actions depicted in Figure~\ref{fig:sequence_actions}. The Emotional Execution Tree for that sequence of actions would be like the one presented in Figure~\ref{fig:emotional_enrichment}.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{./Images/Representation.png}
	\caption{Emotional Execution Tree for the sequence of actions depicted in Figure~\ref{fig:sequence_actions}. The $*$ symbolize that node is principal. S. represents sequential.}
	\label{fig:emotional_enrichment}
\end{figure}

\subsection{Formalization of Simple and Compound Actions}
Formalizing simple and compound actions generates a common system to design them and verify if they are described correctly. Simple actions ($SA$) could be seen as functions that map a set of parameters $p^{pl}_{i} \in P$ to specific platform ($pl$) movements ($m^{pl}_{j}\in M^{pl}$) in the set of movements for that platform ($m^{pl}_{j} =sa^{pl}(\lbrace p^{pl}_{i} \rbrace)$). Each $sa^{pl}$ could have different implementations due to different reasons such as: platform's type, action's goal, how the mapping from inputs to outputs are done, among others. 
In our case, we consider that two actions are different if $(\forall sa_{i}, sa_{j} \in SA, \forall q, t \in PL | Param( sa_{i}^q ) \neq Param( sa_{j}^t )\Rightarrow sa_{i}^q \neq sa_{j}^t)$, 
where $PL$ is the set of all the possible platforms, $Param(\cdot)$ returns the set of parameters that are input of the $sa$ (e.g., $vel$,$pitch$,$text$). 
On the other hand, we will consider two actions as equivalents if $(\forall sa_{i}, sa_{j} \in SA , \forall q, t \in PL | Param( sa_{i}^q ) = Param(sa_{j}^t ) \wedge Obj(sa_{i}^q) = Obj(sa_{j}^t) \Rightarrow sa_{i}^q \equiv sa_{j}^t)$,  
where $Obj(\cdot)$ returns the set of objectives that are intended to be achieved by a given action. This $SA$ equivalence allows us to avoid the specification of all the platforms for all the actions that share the same objective and hide leaf nodes that do not give relevant information to understand the emotional tree.

With this $SA$ definition, it is now possible to define emotion enrichment as $ (\forall sa \in SA, \forall e \in E, \exists \; en$ $|$ $en=Enrichment(sa,e,Param(sa), Intensity(e)))$, where $Intensity(\cdot)$ returns the intensity of a given emotion, $E$ is the set of emotions, and $en$ can be the same $sa$ but with the parameters modified in order to convey the desire emotion $e$, or even a set of $sa$ with their parameters modified. This definition goes along with the idea that new $sa$ could be added to convey a specific emotion.  In addition, the $Enrichment(\cdot)$ function has the following properties:
\begin{itemize}
	\item $(\forall sa_{i},sa_{j} \in SA, \forall e \in E$ $|$ $sa_{i} \equiv sa_{j} \Rightarrow Enrichment(sa_{i},e,...) \equiv Enrichment(sa_{j},e,...))$
	\item $(\forall sa_{i},sa_{j} \in SA, \forall e \in E$ $|$ $(sa_{i} \not\equiv sa_{j}) \Rightarrow Enrichment(sa_{i},e,...) \not \equiv Enrichment(sa_{j},e,...))$
\end{itemize}

The concept of compound action ($CA$) is defined as $g = ca(\{p_{i}\})$, where $(p_{i} \in Param(sa))$, $(sa \subseteq SA)$, $(g \in EXT)$.

### 2017-03-24 RO-MAN-2017-2/Parts/Implementation.tex
Changes in some of the files. Improvements in explanations that were missing before. Improvement of the EBFN description. Shift of the section description after the introduction of the emotional execution three to improve the readability of the description section. I still need to correct and extend some explanations in the article.

The system was implemented in \textit{C++} and interfaced with ROS. The design (Fig.~\ref{fig:system_architecture}) was created following the description presented in previous sections. 
\begin{figure*}
	\centering
	\includegraphics[width=1.0\textwidth]{Images/SystemArchitecture.png} 	
	\caption{General system design. Each simple action corresponds to one ROS node, and there is just one node for the emotion enrichment system. The ovals represent the ROS topic parameters, rectangles represent \textit{black boxes}, and texts outside containers represent input files that contain the system parametrization.}
	\label{fig:system_architecture}
\end{figure*}
The emotion enrichment core is divided in three different modules. 
Each module is responsible for one of the following phases:
\begin{enumerate}
	\item \textit{Generation of emotional execution tree:} this phase starts every time that a new action message is received. The process begins by parsing the format, verifying that the actions described on it exist in the system, and that the parameters correspond to the ones expected by each action included in the message. This parameters' verification is done on the implemented description for each action, which describes the parameters that are mandatory and those that are optional. %TODO????
When the verification is done, and all the actions exist and the parameters correspond, an $ext$ %TODO WHat is this?
is created such as the one presented in Figure~\ref{fig:reference}, which corresponds to the example of walk and speak.
	\begin{figure}
		\centering
		%Representation
	\includegraphics[width=0.45\textwidth]{./Images/Representation.png}
	\caption{Emotional Execution Tree for the example shown in the Figure~\ref{fig:compound_example}. The context node colored in purple represents a sequential context, while the other one represents a parallel.} %TODO Up to now no context node was mentioned. What is it?
	\label{fig:reference}
	\end{figure}
	\item \textit{Emotion addition:} uses the $ext$ created in the previous phase. In this phase new $sa$ %TODO What is an sa??
are added to the $ext$, and the $sa$'s parameters are modified following the emotion description, which is loaded from files. This process is broken down in two steps. First, all the actions that are required to convey the desired emotion, and that are not yet present are added. Second, the emotional parameters are modulated basing on the emotion's intensity and character traits.
	\item \textit{Execution:} this is the last phase and it is done after the $ext$ is ''coloured'' with emotional characteristics (actions additions and emotional parameters). The decision to have two different communication channels, one for action parameters and another for the action emotional parameters, was taken to enable the possibility to update the emotional parameters without interfering with the current execution. In this phase a reference to the mandatory %TODO What is a mandatory action?
and emotional action is maintained, thus, when a new emotion is to be expressed, the system stops the previous emotional action and starts executing the new ones without affecting the general execution of the actions.
\end{enumerate}
%The first one is in charge to receive any action and built a $EXT$. Once this tree is created, the second part is in charge to decorate this tree with the additional actions and modify the parameters of all the actions. The last part is the one that executes the final tree and communicate to all the simple actions ROS nodes the parameters that should be executed by them. It is important to notice that the action's parameters and the emotional parameters are communicated via two different channels. This is done to not affect actions' execution when the emotion or its intensity is changed.
All the text message broadcast among the nodes are written using JSON format, which is human understandable, light, and there are many available parsers.\\ 
To test the system two different platforms were used: Keepon Pro %TODO da citare la prima volta il paper descrittivo
and an own made platform called Triskarino.
Keepon has been used to test the interoperability of the system to different platforms, while Triskarino has been used as a part of the robot actor project~\cite{angel2013}. %TODO ???? The whole TheatreBot architecture has not finished just, thus the change of emotion and selection of action to execute were done manually.
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Keepon Test}
To test the system with Keepon (Fig.~\ref{fig:keepon}) was just necessary to implement the platform's controllers in each one of the simple action ROS nodes. Given that this platform does not have the capability to displace its body, the action move body was not implemented. Once added these controllers to the system, we proceeded to modify the configuration files to use this platform instead of Triskarino. With this small modification, we were able to change from one platform to other. In the video\footnote{YouTube video name Emotional Enrichment System, url: https://youtu.be/bRSXQ0rzkO8}
is shown the test where Keepon had to move its torso forward and backward to express a ``happy'' emotion. 
\begin{figure}
	\centering
	\includegraphics[width=0.2\textwidth]{./Images/Keepon.jpg}
	\caption{Keepon platform.}
	\label{fig:keepon}
\end{figure} 
The action is given by console telling the robot to bend the torso to a desired angle in x. The torso oscillation in y is added automatically by the system following the description given to happiness. 
 %%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Triskarino Test}
The system has been widely used with Triskarino (Fig.~\ref{fig:robot}) during our studies on projecting emotions with a non human like platform. 
\begin{figure}
	\centering
	\begin{subfigure}[c]{0.2\textwidth}
	\includegraphics[width=\textwidth]{./Images/upperFourthD.png}
	\caption{Triskarino design.}
	\label{fig:design}
	\end{subfigure}
	\begin{subfigure}[c]{0.2\textwidth}
	\includegraphics[width=\textwidth]{./Images/platform_fome.jpg}
	\caption{Triskarino with its cover.}
	\label{fig:triskar}
	\end{subfigure}
	\caption{Triskarino platform.}
	\label{fig:robot}
\end{figure}  
 During these studies the system was used in different experiments to move the robot on a straight line showing different emotions, each one selected from our command console. Moreover, to show the whole capabilities of the emotional enrichment system, a little scene was set up. The scene is a modification of the first part of the balcony scene of Shakespeare's Romeo and Juliet play~\cite{RAndJ}. The simplified sketch of our version could be seen in the Fig.~\ref{fig:triskar_test}. As it can be seen the stage was divided in 81 squares and the positions were given to the system in terms of the desire square to be reached, which allows the robot to adjust to the stage's dimension. %TODO I cannot understand what follows. What is a context?
The whole sequence of actions were specified as unique action, having several sequential context and just move body action. The other actions, such as oscillate body or blend upper body were added online accordingly the desired emotion.  
\begin{figure}
	\centering
	%\begin{subfigure}[c]{0.2\textwidth}
	%\includegraphics[width=\textwidth]{./Images/FourthCaseScene.png}
	%\caption{Division of the stage.}  
	%\label{fig:design}
	%\end{subfigure}
	%\begin{subfigure}[c]{0.2\textwidth}
	\includegraphics[width=0.4\textwidth]{./Images/FourthCaseSceneB.png}
	%\caption{Movements.}
	%\label{fig:triskar}
	%\end{subfigure}
	\caption{Sketch of Romeo's movements for the balcony scene in Romeo and Juliet play. The arrows show the direction of the movements, and the numbers their sequence.}
	\label{fig:triskar_test}
\end{figure} 
%The sequence of movements done by the robot could be seen in Fig.~\ref{fig:triskar_movements} and in the video. As it could be seen the robot  and they were selected and executed autonomously by the system.
%\begin{figure*}
%	\centering
%	\includegraphics[width=1.0\textwidth]{Images/Sequence_Triskar.png} 
%	\caption{The actual movements of Triskarino in the Balcony scene.}
%	\label{fig:triskar_movements}
%\end{figure*}

>>>>

%TODO in this section is missing a review on the terms and explain some concepts that are depicted in the figures.

The system was implemented in \textit{C++} and interfaced with ROS. The design (Fig.~\ref{fig:system_architecture}) was created following the description presented in previous sections. 
\begin{figure*}
	\centering
	\includegraphics[width=1.0\textwidth]{Images/SystemArchitecture.png} 	
	\caption{General system design. Each simple action corresponds to one ROS node, and there is just one node for the emotion enrichment system. The ovals represent the ROS topic parameters, rectangles represent \textit{black boxes}, and texts outside containers represent input files that contain the system parametrization.}
	\label{fig:system_architecture}
\end{figure*}
The emotion enrichment core is divided in three different modules. 
Each module is responsible for one of the following phases:
\begin{enumerate}
	\item \textit{Generation of emotional execution tree:} this phase starts every time that a new action message is received. The process begins by parsing the format, verifying that the actions described on it exist in the system, and that the parameters correspond to the ones expected by each action included in the message. This parameters' verification is done on the implemented description for each action, which describes the parameters that are mandatory and those that are optional. %TODO????
When the verification is done, and all the actions exist and the parameters correspond, an Emotional Execution Tree (EXT)
is created such as the one presented in Figure~\ref{fig:emotional_enrichment}
	\item \textit{Emotion addition:} uses the EXT created in the previous phase. In this phase new simple actions ($sa$)
are added to the EXT, and the $sa$'s parameters are modified following the emotion description, which is loaded from files. This process is broken down in two steps. First, all the actions that are required to convey the desired emotion, and that are not yet present are added. Second, the emotional parameters are modulated basing on the emotion's intensity and character traits. The final EXT is presented in Figure~\ref{fig:reference}.
	\begin{figure}
		\centering
		%Representation
	\includegraphics[width=0.45\textwidth]{./Images/exampleTreeE.png}
	\caption{Enrich Emotional Execution Tree.} 
	\label{fig:reference}
	\end{figure}
	\item \textit{Execution:} this is the last phase and it is done after the $ext$ is ''coloured'' with emotional characteristics (actions additions and emotional parameters). The decision to have two different communication channels, one for action parameters and another for the action emotional parameters, was taken to enable the possibility to update the emotional parameters without interfering with the current execution. In this phase a reference to the principal and emotional action is maintained, thus, when a new emotion is to be expressed, the system stops the previous emotional action and starts executing the new ones without affecting the general execution of the actions.
\end{enumerate}

All the text message broadcast among the nodes are written using JSON format, which is human understandable, light, and there are many available parsers.\\ 
To test the system two different platforms were used: Keepon Pro %TODO da citare la prima volta il paper descrittivo
and an own made platform called Triskarino.
Keepon has been used to test the interoperability of the system to different platforms, while Triskarino has been used in diverse case studies to explore emotion projection with a non anthropomorphic platform.
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Keepon Test}
To test the system with Keepon (Fig.~\ref{fig:keepon}) was just necessary to implement the platform's controllers in each one of the simple action ROS nodes. Given that this platform does not have the capability to displace its body, the action move body was not implemented. Once added these controllers to the system, we proceeded to modify the configuration files to use this platform instead of Triskarino. With this small modification, we were able to change from one platform to other. In the video\footnote{YouTube video name Emotional Enrichment System, url: https://youtu.be/bRSXQ0rzkO8}
is shown the test where Keepon had to move its torso forward and backward to express a ``happy'' emotion. 
\begin{figure}
	\centering
	\includegraphics[width=0.2\textwidth]{./Images/Keepon.jpg}
	\caption{Keepon platform.}
	\label{fig:keepon}
\end{figure} 
The action is given by console telling the robot to bend the torso to a desired angle in x. The torso oscillation in y is added automatically by the system following the description given to happiness. 
 %%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Triskarino Test}
The system has been widely used with Triskarino (Fig.~\ref{fig:robot}) during our studies on projecting emotions with a non human like platform. 
\begin{figure}
	\centering
	\begin{subfigure}[c]{0.2\textwidth}
	\includegraphics[width=\textwidth]{./Images/upperFourthD.png}
	\caption{Triskarino design.}
	\label{fig:design}
	\end{subfigure}
	\begin{subfigure}[c]{0.2\textwidth}
	\includegraphics[width=\textwidth]{./Images/platform_fome.jpg}
	\caption{Triskarino with its cover.}
	\label{fig:triskar}
	\end{subfigure}
	\caption{Triskarino platform.}
	\label{fig:robot}
\end{figure}  
 During these studies the system was used in different experiments to move the robot on a straight line showing different emotions, each one selected from our command console. Moreover, to show the whole capabilities of the emotional enrichment system, a little scene was set up. The scene is a modification of the first part of the balcony scene of Shakespeare's Romeo and Juliet play~\cite{RAndJ}. The simplified sketch of our version could be seen in the Fig.~\ref{fig:triskar_test}. As it can be seen the stage was divided in 81 squares and the positions were given to the system in terms of the desire square to be reached, which allows the robot to adjust to the stage's dimension. %TODO I cannot understand what follows. What is a context?
The whole sequence of actions were specified as unique action, having several sequential context and just move body action. The other actions, such as oscillate body or blend upper body were added online accordingly the desired emotion.

### 2017-03-24 RO-MAN-2017-2/Parts/Introduction.tex
Changes in some of the files. Improvements in explanations that were missing before. Improvement of the EBFN description. Shift of the section description after the introduction of the emotional execution three to improve the readability of the description section. I still need to correct and extend some explanations in the article.

The rest of the paper is organized as follows. Section~\ref{sec:related_work} provides a brief overview of particularly relevant work related to our system. Section~\ref{ref:theatre} gives a brief explanation on TheatreBot architecture. Section~\ref{ref:general_system} gives a general introduction to the system ideas. Section~\ref{sec:emotional_execution_tree} gives the basic formalization of our system and principal components terms used on it. Section~\ref{sec:implementation} describes the implementation of the system and shows two demonstrations done with the system using platforms with different capabilities.

>>>>

The rest of the paper is organized as follows. Section~\ref{sec:related_work} provides a brief overview of relevant work. Section~\ref{sec:concepts} introduces all the concepts that are used in the system. Section~\ref{sec:emotional_execution_tree} gives the basic formalization of our system and principal components terms used on it. Section~\ref{sec:description} presents the EBFN created to specify the files used by the system. Section~\ref{sec:implementation} describes the implementation of the system and shows two demonstrations done with the system using platforms with different capabilities.

### 2017-03-24 RO-MAN-2017-2/Parts/RelatedWork.tex
Changes in some of the files. Improvements in explanations that were missing before. Improvement of the EBFN description. Shift of the section description after the introduction of the emotional execution three to improve the readability of the description section. I still need to correct and extend some explanations in the article.

On the other hand, studies focused on entertainment robotics have tried to introduce emotional actions to improve the audience's experience. Breazeal and collaborators~\cite{Breazeal2003} used one robot on the stage. This anemone-like robot had few behaviours, which included getting scared when a person comes too close; the robot was able to show some basic emotions (i.e., fear and interest). Knight~\cite{Knight2011b,Knight2010} used the platform NAO to produce a sort of stand-up comedy. The robot performs basic actions to add some expressiveness to the joke, but it is not intended to project any emotion. Trying to add some theatrical realism, Breazeal and collaborators~\cite{Breazeal2008} designed and implemented a system to control a lamp. The main characteristic of this lamp is that it could be controlled by just one person, by selecting pre-coded emotions.

Other works in performance robotics have developed systems that do not convey any kind of emotion as \textit{Roboscopie}~\cite{Roboscopie2012,Lemaignan2012}, Fan and collaborators~\cite{Fan2009,Fan2013}, and adaptation of Shakespeare's  Midsummer Night's Dream~\cite{murphy2011} use robots in performances with real actors, but without any emotional expression.\\
Although these works use emotions, their main focus was the interpretation of postures or just the use of emotions to increase their robot appealing, but none of them considered the importance of an emotional system that could be used by others. As a consequence, most of these works have created emotional systems that could just work in their specific framework.

>>>>

On the other hand, studies focused on entertainment robotics have tried to introduce emotional actions to improve the audience's experience. Breazeal and collaborators~\cite{Breazeal2003} used one robot on the stage. This anemone-like robot had few behaviours, which included getting scared when a person comes too close; the robot was able to show some basic emotions (i.e., fear and interest). Knight~\cite{Knight2011b} used the platform NAO to produce a sort of stand-up comedy. The robot performs basic actions to add some expressiveness to the joke, but it is not intended to project any emotion. Trying to add some theatrical realism, Breazeal and collaborators~\cite{Breazeal2008} designed and implemented a system to control a lamp. The main characteristic of this lamp is that it could be controlled by just one person, by selecting pre-coded emotions.

Other works in performance robotics have developed systems that do not convey any kind of emotion as \textit{Roboscopie}~\cite{Roboscopie2012}, Fan and collaborators~\cite{Fan2009}, and adaptation of Shakespeare's  Midsummer Night's Dream~\cite{murphy2011} use robots in performances with real actors, but without any emotional expression. Although these works have expressed emotions in their studies, they were not focus on create a framework that could be used by others researchers in HRI. Therefore, the use of affective architectures for virtual agents (e.g. SAIBA/BML framework~\cite{Kopp2006}) could be though as a feasible solution. However these framework describe high level actions without given relevant details to its implementation in physical platform. This gap cannot be closed by making a direct map the abstract actions and the physical platforms. This is due the fact that virtual agents do not have physical constraints, while robots are constrained by their physical capabilities~\cite{Saerbeck2007,Canamero2010}.

### 2017-03-24 RO-MAN-2017-2/Parts/SimpleActions.tex
Changes in some of the files. Improvements in explanations that were missing before. Improvement of the EBFN description. Shift of the section description after the introduction of the emotional execution three to improve the readability of the description section. I still need to correct and extend some explanations in the article.

In order to generate a system that could be used in diverse platforms, an abstraction level in which all of them could fall is required. Therefore simple and compound actions are used to achieve this goal. Simple actions are actions that are considered as primitives: they are used as building blocks. Therefore, these actions are described in the system and are the ones in which the emotional enrichment takes place. Their description specifies mandatory and optional parameters that are required to execute an action. Compound actions are actions that are created from simple actions. These actions are not implemented in the system, but, if it is needed, they can be described in it (e.g., compound actions that are used often).

The simple actions to be implemented to test the system were selected by considering platforms' capabilities and the requirements. The eight actions selected are: move body, oscillate body, move shoulder, oscillate shoulder, move torso, oscillate torso, and do nothing. Description for each action, its mandatory parameters and optional parameters are shown in Table~\ref{table:actions_implemented}.

@@

%TODO add an example for this, very simple and presenting how to do it. IT WOULD BE VERY NICE

>>>>

In order to generate a system that could be used in diverse platforms, an abstraction level in which all of them could fall is required. Therefore simple and compound actions are used to achieve this goal. Simple actions are actions that are considered as primitives: they are used as building blocks. As a consequence, these actions are described in the system and are the ones in which the parameters are changed to embed the emotions. Their description specifies mandatory and optional parameters that are required to execute an action. These parameters are modified accordingly to the information provided in the emotional parameters. Compound actions are actions that are created from simple actions and other compound actions. These actions are not implemented in the system, but, if it is needed, they can be described in it (e.g., compound actions that are used often).

A set of simple actions to be implemented were selected to test the system. These actions were selected considering the capabilities of Keepon and Triskarino. The eight actions selected were: move body, oscillate body, move shoulder, oscillate shoulder, move torso, oscillate torso, and do nothing. Description for each action, its mandatory parameters and optional parameters are presented in Table~\ref{table:actions_implemented}.

@@

To exemplify the use of these actions, lets consider a sequence of movements illustrated in Figure~\ref{fig:triskar_test}. This sequence of movements correspond to an adaptation of the first part of the balcony scene of Shakespeare's Romeo and Juliet play~\cite{RAndJ}. The fist part of this scene is characterized by strong changes in Romeo's emotional state, which is the perfect test for the system.  
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{./Images/FourthCaseSceneB.png}
	\caption{Sketch of Romeo's movements for the balcony scene in Romeo and Juliet play. The arrows show the direction of the movements, and the numbers their sequence.}
	\label{fig:triskar_test}
\end{figure} 
However, just body movements would not exploit the other movements of the platform neither test the whole system. Thus, additional movements are added to illustrate and test the system. A partial view of the whole sequence of actions is illustrated in Figure~\ref{fig:sequence_actions}. The first node correspond to the first action in Figure~\ref{fig:triskar_test}. The second actions corresponds to a compound action created by the parallel execution of the simple action move body and other compound action.

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{./Images/sequenceActions.png}
	\caption{Sequence of simple actions for the movements depicted in Figure~\ref{fig:triskar_test}.}
	\label{fig:sequence_actions}
\end{figure}

### 2017-03-24 RO-MAN-2017-2/Parts/description.tex
Changes in some of the files. Improvements in explanations that were missing before. Improvement of the EBFN description. Shift of the section description after the introduction of the emotional execution three to improve the readability of the description section. I still need to correct and extend some explanations in the article.

The description files allow the parametrization of the system and its adaptation to different circumstances.  The system has the following parametric data:

\begin{itemize}

	\item \textit{Emotion description} gives information of the parameters that should be changed in all the simple actions to express the desired emotion. Therefore, for each combination emotion-action should be given a description of how the parameters are defined. If there is any action that does not have any specification, the system will not change its parameters. The following is the EBFN generated to modify movement parameters: 
\begin{grammar}
[(colon){$\models$}]
[(semicolon)$|$]
[(comma){}]
[(period){\\}]
[(quote){\begin{bf}}{\end{bf}}]
[(nonterminal){$\langle$}{$\rangle$}]
<emotion description>: "'\{'","'emotion:'",<string> "','","'observation:'",<string> "','",<action>"'(,'",<action>)* "'\}'".
<action>:<string>"':\{'",<description> "','",<actions affected> "'\}'".
<description>:"'description:\{'",<action name>"','",<emotion>"','"\\
<parameter's type>"'\}'".
<action name>:"'emotionProfileAction:'",<string>.
<emotion>:"'emotionProfileEmotion:'",<string>.
<parameter's type>: "'movement\_parameter'".
<actions affected>:"'actions:\{'",<action parameter>,\\("','"<action parameter>)*"'\}'".
<action parameter>: <string>"':\{'",<reference>"','"\\<repetition>"','",<parameters>"'\}'".
<reference>: "'reference:'",<number>.
<repetition>: "'repetition:'","'yes'";"'no'".
<parameters>: "'parameters:['" <parameter description>\\("','",<parameter description>)* "']'".
<parameter description>: <movement parameter description>;<new parameters>.
<movement parameter description>:"'\{time:'",<number>"',space:'"\\<number>"'\}'".
\end{grammar}

	\item \textit{Character's emotions} parameters give the system information about character's ''rhythm'' for each pair action-emotion. In other words, it is going to say how the emotional parameters for the desired action should be modulated. If there is not any information, the system will use the default emotional values for the action. The following is the corresponding EBFN:
\begin{grammar}
[(colon){$\models$}]
[(semicolon)$|$]
[(comma){}]
[(period){\\}]
[(quote){\begin{bf}}{\end{bf}}]
[(nonterminal){$\langle$}{$\rangle$}]
<character emotion profile>:"'\{'" <emotion>"','",<actions>"'\}'".
<emotion>: "'emotion:'",<string>.
<actions>: "'actions:['", <action>("','",<actions descriptor>)* "']'".
<action>: "'\{ action: '", <string> "','","'bias:'", <number> "','"\\" 'amplitude:'", <number> ", 'long:'",<number> "\}". 
\end{grammar}

	\item \textit{Action message} contains all the necessary information for to execute an action. The following is a script's EBNF:
\begin{grammar}
[(colon){$\models$}]
[(semicolon)$|$]
[(comma){}]
[(period){\\}]
[(quote){\begin{bf}}{\end{bf}}]
[(nonterminal){$\langle$}{$\rangle$}]
<action>: <simple action>;<compound action>;<context>.
<simple action>: "'\{'" <action header> "','","'parameters:['" <simple action parameters> "']\}'".
<compound action>: "'\{'" <action header> "','","'parameters:['" <compound action parameters> "']\}'".
<action header>: "'type:'", <action type> "','","'name: '" <string>  "','", <is primary>.
<simple action parameters>: "'\{'"<parameter header> "','" ,<parameter description> "'\}'". 
<compound action parameters>:  <simple action parameters> ("','", <simple action parameters>)* .
<context>:"'\{'" <action type> "','", <emotion sync> "','", <action sync> "','", <is primary> "','"\\" 'information:'", <string> "','"," 'actions: '", <action> "'\}'".
<parameter header>: "'type:'", <parameter type> "','"," 'name:'",<string>.
<parameter description>: <parameter amplitude>;<parameter circle>;<parameter landmark>;\\<parameter point>;<parameter speech>;\\<parameter square>;<parameter time>.
<parameter type>: "'mandatory\_parameter'";"'optional\_parameter'". 
<is primary>: "'yes'";"'no'".
<emotion sync>: "'yes'";"'no'".
<action type>: "'parallel\_context'";"'serial\_context'";"'simple\_action'";\\"'composite\_action'".
\end{grammar}
\end{itemize}

>>>>

The description files allow the parametrization and extent the system to fit other applications is required. To standarized the mechanisms to describe these files, it was decided to create a language to describe them. This allows other researchers to create their own files and extended the language to fit their needs. It was decided to use EBNF to describe language.
\subsection{Emotion Description} 
The language created to describe emotion descriptors is the following:
\begin{small}

\begingrammar
%		
<emotion description>: "'\{'" "'emotion:'" <string>"','" "'observation:'" <string>"','" <action> ("','"<action>)* "'\}'".

<action>:<string>"': \{'" <description>"','" <actions affected> "'\}'".

<description>: "'description: \{'" <action name>"','";<emotion>"','" <parameter's type> "'\}'".

<action name>: "'emotionProfileAction:'" <string>.

<emotion>: "'emotionProfileEmotion:'" <string>.

<parameter's type>: "'movement\_parameter'".

<actions affected>: "'actions: \{'" <action parameter> ; ("','" <action parameter>)* "'\}'".

<action parameter>: <string>"':\{'" <reference>"','" <repetition>"','" <parameters> "'\}'".

<reference>: "'reference:'" <number>.

<repetition>: "'repetition:'" "'yes'" | "'no'".

<parameters>: "'parameters: ['" <parameter description> ("','" <parameter description>)* "']'".

<parameter description>: <movement parameter description>.

<movement parameter description>: "'\{time:'" <number> "', space:'" <number> "'\}'".

\endgrammar
\end{small}
%TODO give information regardin 1) why repetition 2) why a set of parameters 3) why use of time and space

\subsection{Action Message}
The language created to describe action Messages is the following:
\begin{small}
\begingrammar
%		
<action>: <simple action> <compound action> <context>.

<simple action>: "'\{'" <action header>"','" "'parameters: ['" <simple action parameters> "'] \}'".

<compound action>: "'\{'" <action header>"','" "'parameters:['" <compound action parameters> "']\}'".

<action header>: "'type:'" <action type>"','" "'name: '" <string>"','" <is primary>.

<simple action parameters>: "'\{'" <parameter header>"','" <parameter description> "'\}'".

<compound action parameters>:  <simple action parameters> ("','" <simple action parameters>)* .

<context>: "'\{'" <action type>"','" <emotion sync>"','" <action sync>"','" <is primary>"','" " 'information:'" <string>"','" " 'actions: '" <action> "'\}'".

<parameter header>: "'type:'" <parameter type>"','" " 'name:'" <string>.

<parameter description>: <parameter amplitude> |; <parameter circle> | <parameter landmark> | <parameter point> | <parameter speech> | <parameter square> | <parameter time>.

<parameter type>: "'mandatory\_parameter'" | ;"'optional\_parameter'".

<is primary>: "'yes'" | "'no'".

<emotion sync>: "'yes'" | "'no'".

<action type>: "'parallel\_node'" | "'sequential\_node'" | ;"'simple\_action'" | "'composite\_action'".

\endgrammar
\end{small}

%TODO give information regarding: primary and emotion synch
